---
title: "Holt-Winters example"
author: "Bernardo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(readr)
library(ggplot2)
library(forecast)
library(stats)

```


```{r cars}
shampoo <- read_csv(file="C:/Users/Bernardo Carvalho/OneDrive - Imperial College London/LCS/shampoo.csv")



ggplot(ts, aes(Month, "Sales of shampoo over a three year period")) + 
  geom_line()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


```{r pressure, echo=FALSE}
file <- "C:/Users/Bernardo Carvalho/OneDrive - Imperial College London/LCS/shampoo.csv"

data <- read.csv(file = file, header = TRUE)

shampoo <- ts(data[, 2], frequency = 12, start = c(2001, 1))

shampoo %>% stl(s.window = "period") %>% autoplot
```
```{r}
shampoo.ets <- ets(shampoo, model = "ZZZ")

```
With the ets() function, the best model is given by model = "AAN", which suggests that the error component is additive, trend is additive, and there is no evidence of seasonality. This is consistent with conclusions from our visual inspection.

```{r}
shampoo.ets.f <- forecast(shampoo.ets, h = 6)
```
After estimation, we can use forecast() function to generate forecast.

```{r}
shampoo.diff1 <- diff(shampoo, differences = 1)
autoplot(shampoo.diff1)

```
```{r}
library(tseries)

adf.test(shampoo.diff1)
pp.test(shampoo.diff1)
kpss.test(shampoo.diff1)
```
```{r}
ndiffs(shampoo)
nsdiffs(shampoo)

```

```{r}
acf(shampoo)
```
```{r}
pacf(shampoo)
```
```{r}
ggAcf(shampoo.diff1)
ggPacf(shampoo.diff1)

```

Based on the plots, we know that the order
 of MA is no greater than 2 (as the first two lags of ACF are significant), and the order of AR is no greater
 than 1 (as only the first lag of PACF is significant).
 
 Next we use auto.arima() to evaluate the quality of various ARIMA models. Potentially, we can use our
 earlier analysis as input to refine the models that will be evaluated by auto.arima(). But for this simple 
exercise, we just use the default setting from auto.arima(), and run a stepwise search.
 
```{r}
auto.arima(shampoo, trace = TRUE, ic = 'bic')
```
In the post-estimation  analysis, we should also check out the residual plots, including the time series and ACFs to make sure 
that there is no warning message. In particular, residuals should have a zero mean, constant variance, and should be distributed symmetrically around mean zero. ACF of any lag greater than zero is statistically insignificant.
shampoo.m1 <- Arima(shampoo, order = c(1, 1, 1),
seasonal = list(order = c(0, 0, 1), period = 12), include.drift = TRUE)

```{r}
shampoo.m1 <- Arima(shampoo, order = c(1, 1, 1), seasonal = list(order = c(0, 0, 1), period = 12), include.drift = TRUE)

autoplot(shampoo.m1$residuals)
ggAcf(shampoo.m1$residuals)
checkresiduals(shampoo.m1)

```

Residuals appear to be random, and there exists no inter-temporal correlation. So there is no warning message. Once we have selected a model, forecasting is easy - it is readily implemented by forecast() function 
from the forecast package.

```{r}
shampoo.f <- forecast(shampoo.m1, h = 6)
autoplot(shampoo.f)

```

 